### 2024-08-20 paper.tex
start rehabilitating normalization proof

\subsection{A `normalization' fact about Jordan bases}
Let \(V\) be a finite-dimensional vector space, \(A : V \to V\) a nilpotent operator, and \(f : V \to \mathbb{C}\) a linear map.
Let \((e_{ij} : i \leq m, j \leq \lambda_i)\) be a Jordan basis for \(A\).

\begin{lemma}
    There is a Jordan basis \(e_{ij}\) for \(A\) such that there is at most one pair \((i, j)\) with \(f(e_{ij}) \neq 0\).
\end{lemma}
\begin{proof}
    For any change of basis \(P : V \to V\) commuting with \(A\), we obtain a new Jordan basis \((P(e_{ij}) : i \leq m, j \leq \lambda_i)\).
    For any such \(P\), define 
    \[S_P = \sum_i \begin{cases}-1, & \forall j. \; f \circ P(e_{ij}) = 0 \\ \lambda_i - \min\{j : f \circ P(e_{ij}) \neq 0\}, & \textrm{otherwise} \end{cases}.\]
    Let \(P\) be any operator, among all invertible operators commuting with \(A\), which minimizes \(S_P\).
    Write \(e_{i,j}' := P(e_{i,j})\).
    \par Suppose for contradiction that there are two distinct \(i\)'s (and some \(j\)'s) with \(f(e_{ij}') \neq 0\).
    Then we can take \(e_{i_1, j_1}'\) and \(e_{i_2, j_2}'\), where for \(k \in \{1,2\}\) we have \(f(e_{i_k, j_k}') \neq 0\), and \(\forall j < j_k. \; f(e_{i_k, j}') = 0\).
    Wlog, assume \(\lambda_{i_1} - j_1 \leq \lambda_{i_2} - j_2\).
    Then, we can define \(Q : V \to V\) by
    \begin{itemize}
        \item \(Q(e_{i_1, \lambda_1}') := e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\)
        \item For \(j < \lambda_1\), \(Q(e_{i_1, j}') := A^{\lambda_{i_1} - j} Q(e_{i_1, \lambda_{i_1}}')\)
        \item For \(i \neq i_1\), \(Q(e_{ij}') = e_{ij}'\).
    \end{itemize}
    Clearly \(Q\) is invertible, and it commutes with \(A\).
    Further, I claim that \(S_{QP}  < S_P\).
    It suffices to show that \(\forall j \leq j_1. \; f \circ Q(e_{i_1,j}') = 0\).
    We have 
    \[f \circ Q(e_{i_1,j}') = f\left(A^{\lambda_{i_1} - j}\left(e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\right)\right) =\]
    \[f\left(e_{i_1, j}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e_{i_2, j_2 + (j - j_1)}'\right) = f(e_{i_1, j}') - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e'_{i_2, j_2 + (j - j_1)}.\]
    Clearly (by design), this expression is zero when \(j = j_1\).
    And for \(j < j_1\), we have \(f(e'_{i_1, j}) = f(e'_{i_2, j_2 + (j_2 - j_1)}) = 0\), so it is zero then as well.
    Thus we see that \(S_{QP} < S_P\), contradicting that \(S_P\) is minimal.
    So there must be at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}') \neq 0\).
    \par If there is no such \(i\), we have found the desired basis.
    So, suppose there is such an \(i_0\).
    Let \(U = \langle e_{i_0, j}' \rangle\).
    Simply write \(e_j := e_{i_0, j}'\).
    Let \(j_0 = \min\{j : f(e_j) \neq 0\}\).
    We just have to change basis to zero out \(e_j\) for \(j \neq j_0\).
    Let \(P_1(e_{\lambda_{i_0}}) := e_{\lambda_{i_0}} - \frac{f(e_{j_0 + 1})}{f(e_{j_0})} e_{\lambda_{i_0} - 1}\), and notice that \(f \circ P_1(e_j) = 0\) for all \(j \leq j_0 + 1\) except for \(j_0\).
    Then we define \(P_2(e_{\lambda_{i_0}}) := P_1(e_{\lambda_{i_0}}) - \frac{f\circ P_1(e_{j_0 + 2})}{f\circ P_1(e_{j_0})} e_{\lambda_{i_0} - 2}\), and notice that \(f \circ P_2(e_j) = 0\) for all \(j \leq j_0 + 2\) except for \(j_0\).    
    Eventually we get \(P_{\lambda_{i_0} - j_0}\), and by applying this to the \(e_j\)'s we obtain a basis of \(U\), in which there is exactly one \(j\) with \(f \circ P_{\lambda_{i_0} - j_0}(e_j) \neq 0\).
\end{proof}

@@

\section{Finding the irreducible components of \(U_0\)}
\subsection{Introduction}
Let \(X\) be nilpotent with Jordan basis \((e_{ij}')_{1 \leq i \leq k, 1 \leq j \leq \lambda_i'}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
For each row \(i\), let \(p_i(b) = \max\{j : b_{ij} \neq 0\}\) (the maximum of the empty set is zero).
Then set \(q_i(b) = \lambda_i - p_i(b)\).
When it is clear enough from context, we will just write \(p_i\) and \(q_i\) instead of \(p_i(b)\) and \(q_i(b)\).
\par For any \(I \subseteq \{1, ..., k\}\) and any values \((\rho_i)_{i \in I}\) %such that (1) \(\rho_i \leq \lambda_i\), (2) \(\rho_i\) is decreasing with \(i\), and (3) \(\lambda_i - \rho_i\) is decreasing with \(i\)
, define 
\[B_{I, (\rho_i)} := \{b \in \mathbb{C}^m : \]
\[\{(p_i, q_i) : i \in I\} = \{(p_i, q_i) : n > q_i \land p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j \} \] 
\[\land \forall i \in I. \; p_i = \rho_i \}.\]
Then let \(U_{I, (\rho_i)} = \{(A_{0, b}, U) \in V : b \in B_{I, (\rho_i)}\}\).
\begin{lemma}
    \(\mathbb{C}^m = \bigcup_{I, (\rho_i)} B_{I, (\rho_i)}\), where \(I\) ranges over all subsets of \(\{1, ..., k\}\), and \((\rho_i)\) ranges over all maps \(I \to \mathbb{N}_{>0}\) such that \(\rho_i\) and \(\lambda_i - \rho_i\) are both decreasing with \(i\).
    TODO: something about overlap being small (they're not disjoint).  
    They are distinct.
    Is that good enough for my purposes?
    Ah, I think it is good enough if I instead say that the sets of \(b\)'s are distinct.
\end{lemma}
\begin{proof}
    Let \(b \in \mathbb{C}^m\).
    Let \(I' = \{(p_i, q_i) : p_i = n > q_i \land \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j\}\).
    Then let \(I = \{\max\{i : (p_i, q_i) = (p, q)\} : (p, q) \in I'\}\).
    Choosing the maximal \(i\) is an arbitrary choice; we just have to pick one.
    For \(i \in I\), \(\rho_i := p_i\).
    All we need to do is show that each \(\rho_i\) is positive, and that \(\rho_i\) and \(\lambda_i - \rho_i\) are decreasing.
    Then we get that \((A_{0,b}, U) \in U_{I, (\rho_i)}\), and we are done.
    \par Recall that by convention the maximum of the empty set is zero.
    So, the fact that \(p_i > \max_{j > i} p_j\) tells us that \(\rho_i = p_i\) is positive.
    \par Let \(i, j \in I\) with \(i < j\).
    Since \(\lambda_i \geq \lambda_j\), we have either \(p_i \geq p_j\) or \(q_i \geq q_j\).
    Further, since \((p_i, q_i) \neq (p_j, q_j)\) (clear from definition of \(I\)), either \(p_i > p_j\) or \(q_i > q_j\).
    And for \(i, j \in I\) we clearly have \(p_i > p_j \iff q_i > q_j\).
    So in either case, both \(p_i > p_j\) and \(q_i > q_j\), and consequently both \(\rho_i > \rho_j\) and \(\lambda_i - \rho_i > \lambda_j - \rho_j\).
\end{proof}

>>>>

\subsection{A `normalization' fact about Jordan bases}
Let \(V\) be a finite-dimensional vector space, \(A : V \to V\) a nilpotent operator, and \(f : V \to \mathbb{C}\) a linear map.
% Let \((e_{ij} : i \leq m, j \leq \lambda_i)\) be a Jordan basis for \(A\).

\begin{lemma}
    There is a Jordan basis \(e_{ij}\) for \(A\) such that there is at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}) \neq 0\).
\end{lemma}
\begin{proof}
    For any Jordan basis \((e_{ij})_{ij}\) of \(A\), define
    \[S((e_{ij})_{1 \leq i \leq k, 1 \leq j \leq \lambda_i}) := \sum_i \begin{cases}-1, & \forall j. \; f (e_{ij}) = 0 \\ \lambda_i - \min\{j : f (e_{ij}) \neq 0\}, & \textrm{otherwise} \end{cases}.\]
    We proceed by induction on the measure \(S\).
    That is, let \((e_{ij})_{ij}\) be a Jordan basis for \(A\).
    Our inductive hypothesis is that if there exists a Jordan basis \((e_{ij}')_{ij}\) with \(S((e_{ij}')_{ij}) < S((e_{ij})_{ij})\), then we get the desired conclusion.
    \par Now, we have two cases.
    In the first case, \((e_{ij})_{ij}\) already satisfies the desired property.
    In this case we are done.
    In the other case, there exist \(i_1, j_1, i_2, j_2\) with \(f(e_{i_1j_1}) \neq 0\) and \(f(e_{i_2j_2}) \neq 0\).
    We let \(j_1, j_2\) be minimal with this property, so that \(\forall j < j_1. \; f(e_{i_1j}) = 0\), and \(\forall j < j_2. \; f(e_{i_2j}) = 0\).
    Wlog, we assume that \(\lambda_{i_1} - j_1 \leq \lambda_{i_2} - j_2\).
    \par By our inductive hypothesis, all we need to do is find a Jordan basis \((e_{ij}')_{ij}\) with \(S((e_{ij}')_{ij}) < S((e_{ij})_{ij})\).
    This is what we do.
    Define \(e'_{ij}\) as follows.
    \begin{itemize}
        \item \(e_{i_1, \lambda_1}' := e_{i_1, \lambda_1} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}\)
        \item CONTINUE HERE
        \item For \(j < \lambda_1\), \(Q(e_{i_1, j}') := A^{\lambda_{i_1} - j} Q(e_{i_1, \lambda_{i_1}}')\)
        \item For \(i \neq i_1\), \(Q(e_{ij}') = e_{ij}'\).
    \end{itemize}
    Clearly \(Q\) is invertible, and it commutes with \(A\).
    Further, I claim that \(S_{QP}  < S_P\).
    It suffices to show that \(\forall j \leq j_1. \; f \circ Q(e_{i_1,j}') = 0\).
    We have 
    \[f \circ Q(e_{i_1,j}') = f\left(A^{\lambda_{i_1} - j}\left(e_{i_1, \lambda_1}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}'\right)\right) =\]
    \[f\left(e_{i_1, j}' - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e_{i_2, j_2 + (j - j_1)}'\right) = f(e_{i_1, j}') - \frac{f(e'_{i_1, j_1})}{f(e'_{i_2, j_2})} e'_{i_2, j_2 + (j - j_1)}.\]
    Clearly (by design), this expression is zero when \(j = j_1\).
    And for \(j < j_1\), we have \(f(e'_{i_1, j}) = f(e'_{i_2, j_2 + (j_2 - j_1)}) = 0\), so it is zero then as well.
    Thus we see that \(S_{QP} < S_P\), contradicting that \(S_P\) is minimal.
    So there must be at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}') \neq 0\).
    \par If there is no such \(i\), we have found the desired basis.
    So, suppose there is such an \(i_0\).
    Let \(U = \langle e_{i_0, j}' \rangle\).
    Simply write \(e_j := e_{i_0, j}'\).
    Let \(j_0 = \min\{j : f(e_j) \neq 0\}\).
    We just have to change basis to zero out \(e_j\) for \(j \neq j_0\).
    Let \(P_1(e_{\lambda_{i_0}}) := e_{\lambda_{i_0}} - \frac{f(e_{j_0 + 1})}{f(e_{j_0})} e_{\lambda_{i_0} - 1}\), and notice that \(f \circ P_1(e_j) = 0\) for all \(j \leq j_0 + 1\) except for \(j_0\).
    Then we define \(P_2(e_{\lambda_{i_0}}) := P_1(e_{\lambda_{i_0}}) - \frac{f\circ P_1(e_{j_0 + 2})}{f\circ P_1(e_{j_0})} e_{\lambda_{i_0} - 2}\), and notice that \(f \circ P_2(e_j) = 0\) for all \(j \leq j_0 + 2\) except for \(j_0\).    
    Eventually we get \(P_{\lambda_{i_0} - j_0}\), and by applying this to the \(e_j\)'s we obtain a basis of \(U\), in which there is exactly one \(j\) with \(f \circ P_{\lambda_{i_0} - j_0}(e_j) \neq 0\).
\end{proof}

@@

\section{Finding the irreducible components of \(U_0\)}
\subsection{Introduction}
Let \(X\) be nilpotent with Jordan basis \((e_{ij}')_{1 \leq i \leq k, 1 \leq j \leq \lambda_i'}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
For each row \(i\), let \(p_i(b) = \max\{j : b_{ij} \neq 0\}\) (the maximum of the empty set is zero).
Then set \(q_i(b) = \lambda_i - p_i(b)\).
When it is clear enough from context, we will just write \(p_i\) and \(q_i\) instead of \(p_i(b)\) and \(q_i(b)\).
\par For any \(I \subseteq \{1, ..., k\}\) and any values \((\rho_i)_{i \in I}\) %such that (1) \(\rho_i \leq \lambda_i\), (2) \(\rho_i\) is decreasing with \(i\), and (3) \(\lambda_i - \rho_i\) is decreasing with \(i\)
, define 
\[B_{I, (\rho_i)} := \{b \in \mathbb{C}^m : \]
\[\{(p_i, q_i) : i \in I\} = \{(p_i, q_i) : n > q_i \land p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j \} \] 
\[\land \forall i \in I. \; p_i = \rho_i \}.\]
Then let \(U_{I, (\rho_i)} = \{(A_{0, b}, U) \in V : b \in B_{I, (\rho_i)}\}\).
\begin{lemma}\label{bs_union}
    \(\mathbb{C}^m = \bigcup_{I, (\rho_i)} B_{I, (\rho_i)}\), where \(I\) ranges over all subsets of \(\{1, ..., k\}\), and \((\rho_i)\) ranges over all maps \(I \to \mathbb{N}_{>0}\) such that \(\rho_i\) and \(\lambda_i - \rho_i\) are both decreasing with \(i\).
\end{lemma}
\begin{proof}
    Let \(b \in \mathbb{C}^m\).
    Let \(I' = \{(p_i, q_i) : p_i = n > q_i \land \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j\}\).
    Then let \(I = \{\max\{i : (p_i, q_i) = (p, q)\} : (p, q) \in I'\}\).
    Choosing the maximal \(i\) is an arbitrary choice; we just have to pick one.
    For \(i \in I\), \(\rho_i := p_i\).
    All we need to do is show that each \(\rho_i\) is positive, and that \(\rho_i\) and \(\lambda_i - \rho_i\) are decreasing.
    Then we get that \((A_{0,b}, U) \in U_{I, (\rho_i)}\), and we are done.
    \par Recall that by convention the maximum of the empty set is zero.
    So, the fact that \(p_i > \max_{j > i} p_j\) tells us that \(\rho_i = p_i\) is positive.
    \par Let \(i, j \in I\) with \(i < j\).
    Since \(\lambda_i \geq \lambda_j\), we have either \(p_i \geq p_j\) or \(q_i \geq q_j\).
    Further, since \((p_i, q_i) \neq (p_j, q_j)\) (clear from definition of \(I\)), either \(p_i > p_j\) or \(q_i > q_j\).
    And for \(i, j \in I\) we clearly have \(p_i > p_j \iff q_i > q_j\).
    So in either case, both \(p_i > p_j\) and \(q_i > q_j\), and consequently both \(\rho_i > \rho_j\) and \(\lambda_i - \rho_i > \lambda_j - \rho_j\).
\end{proof}

\begin{lemma}
    There are blah \(B_{I, (\rho_i)}\) with \((\rho_i)\) satisfying the constraints of \cref{bs_union}, and none is contained in the union of the others.
\end{lemma}
\begin{proof}
    seems hard to count.
    total number is product over L of (number of rows of length L) times (count where we assume there is only one row of each length).
